---
title: "CreativeAssignmentTwo"
author: "Whytne Stevens"
date: "9/18/2020"
output: html_document
---
# Creative Assignment Two

For this creative assignment, I was itnersted in looking at urban areas on the African continent
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
```


First I pull a world map using the libraries loaded above.
```{r}
world <- ne_countries(scale = 50, returnclass = "sf") 

world_map <- ggplot(world) +
  geom_sf() +
  theme_solarized_2()

world_map
```

I then pulled a map of the African continent with country boundaries drawn on the map
```{r}
africa_countries <- world %>%
  filter(continent == "Africa")

africa_map <- ggplot(africa_countries) +
  geom_sf() + 
  theme_solarized_2()

africa_map
```

```{r}
africa_map +
  annotation_scale()
```
Next I defined my coordinate systems. I decided on one that was more___(kklgk), one regional, and one local.

```{r define crs options}
Africa_AEA_Conic <- "+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
Abidjan_1987 <- "+proj=longlat +ellps=clrk80 +towgs84=-124.76,53,466.79,0,0,0,0 +no_defs"
East_Africa <- "+proj=utm +zone=36 +south +ellps=clrk80 +units=m +no_defs"
North_African_Bonne <- "+proj=bonne +lon_0=2.337208 +lat_1=35.1 +x_0=0 +y_0=0 +a=6378249.2 +b=6356514.999904194 +units=m +no_defs"
Test_French_West_Africa_Niger_zone <- "+proj=tmerc +lat_0=0 +lon_0=7.3 +k=0.999 +x_0=1000000 +y_0=1000000 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
Africa_Sinusoidal <- "+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
Africa_Equidistant_Conic <- "+proj=eqdc +lat_0=0 +lon_0=0 +lat_1=20 +lat_2=-23 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs "
Africa_Lambert_Conformal_Conic <- "+proj=lcc +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
brads <- "+proj=tmerc +lat_0=0 +lon_0=31 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"
```

```{r}
africa_map +
  geom_sf() +
  coord_sf(crs = Africa_Sinusoidal) +
  theme_solarized_2()
```

```{r}
africa_map +
  geom_sf() +
  coord_sf(crs = North_African_Bonne) +
  theme_solarized_2()
```

```{r}
africa_map +
  geom_sf() +
  coord_sf(crs = Test_French_West_Africa_Niger_zone) +
  theme_solarized_2()
```

```{r}
africa_map +
  geom_sf() +
  coord_sf(crs = East_Africa) +
  theme_solarized_2()
```
Having pulled these projections, I then pulled data on the urban areas in Africa

```{r}
urbanareas50 <- ne_download(scale = 50, type = 'urban_areas', category = 'cultural')
sp::plot(urbanareas50)
```

```{r}
urbanareas10 <- ne_download(scale = 10, type = 'urban_areas', category = 'cultural')
sp::plot(urbanareas10)
```









